---
title: "Evaluate best model for detection"
author: Florian Tanner
date: "`r format(Sys.time())`"
output: html_document
---


```{r}
.libPaths("C:/R-packages2")
rm(list=ls())
```

## Content 

Fit to entire 2021 data (2DAI best model as evaluated in meta_analysis_detect_ab.Rmd)
Apply to 2020 data
Extract important wavelengths
2	Full spectrum	Raw	VNIR + SWIR	Pu	First	SVM

## Load packages

```{r,  message = FALSE}
library(tidyverse)
library(ggplot2); theme_set(theme_bw())
library(tidymodels)
library(prospectr)
library(gt)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data

```{r}
hyperspec_full_2020 <- read_csv("../pre_processing/hyperspectral/out/2020/raw_and_smoothed_full.csv")
hyperspec_full_2021 <- read_csv("../pre_processing/hyperspectral/out/2021/raw_and_smoothed_full.csv")
hyperspec_FWHM_2020 <- read_csv("../pre_processing/hyperspectral/out/2020/FWHM_binned.csv")
hyperspec_FWHM_2021 <- read_csv("../pre_processing/hyperspectral/out/2021/FWHM_binned.csv")
hyperspec_double_FWHM_2020 <- read_csv("../pre_processing/hyperspectral/out/2020/double_FWHM_binned.csv")
hyperspec_double_FWHM_2021 <- read_csv("../pre_processing/hyperspectral/out/2021/double_FWHM_binned.csv")
scores_2020 <- read_csv("../raw_data/scores/scores_2020.csv") 
scores_2021 <- read_csv("../raw_data/scores/scores_2021.csv") 
```

### Function to prepare dataset for modeling


```{r}
prep_dataset <- function(dataset = hyperspec_full, 
                         alive_only = TRUE, 
                         response_chr = "reflectance_raw", 
                         VNIR_only = TRUE, 
                         set_date = as.Date("2021-06-15"),
                         normalization = "Pu", 
                         derivative = "None",
                         gapDer_wl = 11,
                         gapDer_seg = 9,
                         scores = scores){
  
  # Combine with scores and pot ids
  dat <- dataset |>
    rename(imaging_date = date) |> # do not use function names as variable names
    left_join(scores) |> 
    dplyr::filter(imaging_date == set_date) |> 
    rename(pot = id_tag)
  
  try(dat <- dat |> rename(wavelength = wavelength_bin)) # Rename var for the binned datasets
  
  # Subset
  if(alive_only == TRUE){
    dat <- dat |> filter(di < 100)
  }else{
    dat <- dat
  }
  if(VNIR_only == TRUE){
    dat <- dat |>  filter(camera == "VNIR")
  }else{
    dat <- dat
  }
  
  # Pre-processing
  
  ## Normalization
  if(normalization == "Pu"){
    pots <- unique(dat$pot)
    pot_summary <- data.frame(pot = character(),
                              avg_ref_per_curve = numeric())
    for(p in pots){ # Summarizing reflectance per pot
      pot_subset <- dat[dat$pot == p,]
      list_of_response <- pot_subset |> pull(response_chr)
      pot_mean <- mean(list_of_response, na.rm = TRUE)
      p_summary <- data.frame(pot = p, avg_ref_per_curve = as.numeric(pot_mean))
      
      pot_summary <- rbind(pot_summary, p_summary)
    }
    dat <- dat |> 
      left_join(pot_summary) 
    response_num <- dat |> pull(response_chr)
    
    dat$response <- response_num/dat$avg_ref_per_curve
  }else if(normalization == "None"){
    dat <- dat
    dat$response <- dat |> pull(response_chr)
  }else{
    dat <- dat
    dat$response <- dat |> pull(response_chr)
  }
  
  
  ## Calculate derivatives 
  
  # Cast wide
  dat <- dat |>  
    pivot_wider(id_cols = c(pot, di), values_from = response, 
                names_from = wavelength,values_fn = mean)
  
  # Calculate lagged Differences (derivatives), adjust lag depending on binning
  if(derivative == "None"){
    dat <- dat
    
  }else if(derivative == "First"){
    hyper_mat <- as.matrix(dat[ , !names(dat) %in% c("pot","di")])
    hyper_id <- as.matrix(dat[ , names(dat) %in% c("pot","di")])
    
    gds1 <- gapDer(X = hyper_mat, m = 1, w = gapDer_wl, s = gapDer_seg)

    dat <- data.frame(hyper_id, gds1, check.names = FALSE)
    dat$di <- as.numeric(dat$di)
    
  }else if(derivative == "Second"){
    hyper_mat <- as.matrix(dat[ , !names(dat) %in% c("pot","di")])
    hyper_id <- as.matrix(dat[ , names(dat) %in% c("pot","di")])
    
    gds2 <- gapDer(X = hyper_mat, m = 2, w = gapDer_wl, s = gapDer_seg)
    
    dat <- data.frame(hyper_id, gds2, check.names = FALSE)
    dat$di <- as.numeric(dat$di)
    
  }else{
    stop('Derivative needs to be "None", "First", or "Second"')
  }
  dat <- dat |> left_join(scores |> select(treatment, id_tag) |> rename(pot = id_tag)) |> select(-di)
  return(dat)
}
```

# Evaluate 2020 algorithms on 2021 data

## Fit final model to 2021 data (choose best one in meta_analysis_detect_ab)

## Prepare data

### 2DAI

```{r}
data_2020_2dai_best_process_2020 <- prep_dataset(dataset = hyperspec_double_FWHM_2020, 
                                       alive_only = TRUE, # Not relevant here, no dead pots
                                       response_chr = "reflectance_raw", 
                                       set_date = as.Date("2020-05-14"),
                                       normalization = "Pu",
                                       VNIR_only = TRUE, 
                                       derivative = "Second",
                                       scores = scores_2020)

data_2021_2dai_best_process_2020 <- prep_dataset(dataset = hyperspec_double_FWHM_2021, 
                                       alive_only = TRUE, # Not relevant here, no dead pots
                                       response_chr = "reflectance_raw", 
                                       set_date = as.Date("2021-05-27"),
                                       normalization = "Pu",
                                       VNIR_only = TRUE, 
                                       derivative = "Second",
                                       scores = scores_2021)

colnames(data_2021_2dai_best_process_2020) <- colnames(data_2020_2dai_best_process_2020)
```

### 6DAI

```{r}
data_2020_6dai_best_process_2020 <- prep_dataset(dataset = hyperspec_full_2020, 
                                       alive_only = TRUE, # Not relevant here, no dead pots
                                       response_chr = "reflectance_raw", 
                                       set_date = as.Date("2020-05-18"),
                                       normalization = "None",
                                       VNIR_only = TRUE, 
                                       derivative = "Second",
                                       scores = scores_2020)

data_2021_6dai_best_process_2020 <- prep_dataset(dataset = hyperspec_full_2021, 
                                       alive_only = TRUE, # Not relevant here, no dead pots
                                       response_chr = "reflectance_raw", 
                                       set_date = as.Date("2021-05-31"),
                                       normalization = "None",
                                       VNIR_only = TRUE, 
                                       derivative = "Second",
                                       scores = scores_2021)


colnames(data_2021_6dai_best_process_2020) <- colnames(data_2020_6dai_best_process_2020)
```

### 13DAI

```{r}
data_2020_13dai_best_process_2020 <- prep_dataset(dataset = hyperspec_FWHM_2020, 
                                       alive_only = TRUE, # Not relevant here, no dead pots
                                       response_chr = "reflectance_smoothed", 
                                       set_date = as.Date("2020-05-25"),
                                       normalization = "Pu",
                                       VNIR_only = FALSE, 
                                       derivative = "None",
                                       scores = scores_2020)

data_2021_13dai_best_process_2020 <- prep_dataset(dataset = hyperspec_FWHM_2021, 
                                       alive_only = TRUE, # Not relevant here, no dead pots
                                       response_chr = "reflectance_smoothed", 
                                       set_date = as.Date("2021-06-07"),
                                       normalization = "Pu",
                                       VNIR_only = FALSE, 
                                       derivative = "None",
                                       scores = scores_2021)



colnames(data_2021_13dai_best_process_2020) <- colnames(data_2020_13dai_best_process_2020)
```

### 21 DAI

```{r}
data_2020_21dai_best_process_2020 <- prep_dataset(dataset = hyperspec_full_2020, 
                                       alive_only = TRUE, # Not relevant here, no dead pots
                                       response_chr = "reflectance_raw", 
                                       set_date = as.Date("2020-06-02"),
                                       normalization = "None",
                                       VNIR_only = TRUE, 
                                       derivative = "Second",
                                       scores = scores_2020)

data_2021_21dai_best_process_2020 <- prep_dataset(dataset = hyperspec_full_2021, 
                                       alive_only = TRUE, # Not relevant here, no dead pots
                                       response_chr = "reflectance_raw", 
                                       set_date = as.Date("2021-06-07"),
                                       normalization = "None",
                                       VNIR_only = TRUE, 
                                       derivative = "Second",
                                       scores = scores_2021)



colnames(data_2021_21dai_best_process_2020) <- colnames(data_2020_21dai_best_process_2020)
```

## Fit model to 2020 data

### Set engine

```{r}
rf_spec <- rand_forest() |>
  set_engine("ranger") |>
  set_mode("classification") |>
  translate()


svm_spec <- svm_rbf(cost = 50) |>
    set_engine("kernlab") |>
    set_mode("classification") |>
    translate()
```


### 2DAI

```{r}
detect_recipe_2DAI_2020 <- recipe((treatment ~ .), 
                        data = data_2020_2dai_best_process_2020)  |>
  update_role(pot, new_role = "ID") |> 
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |>
  step_zv(all_predictors()) 

rf_wf_2DAI_2020 <- workflow() |>
  add_model(rf_spec) |>
  add_recipe(detect_recipe_2DAI_2020)

set.seed(123)
rf_fit_2020_2DAI_2020 <- rf_wf_2DAI_2020 |> fit(data_2020_2dai_best_process_2020)
```

### Evaluate on 2021 data

```{r}
preds_2021_2dai <- predict(rf_fit_2020_2DAI_2020, data_2021_2dai_best_process_2020) |> 
  bind_cols(data_2021_2dai_best_process_2020$treatment) |> 
  mutate(truth = as.factor(...2)) 

preds_2021_2dai |> 
  accuracy(truth = truth, estimate = .pred_class)

conf_mat_2021_2dai <- preds_2021_2dai %>% 
  conf_mat(truth = truth, estimate = .pred_class) %>% 
  tidy() %>% 
  mutate(prediction = rep(c("Infected", "Control"), 2),
         truth = c(rep("Infected", 2), rep("Control", 2)),
         predicted_year = 2021,
         predicted_dai= 2)
```

### 6DAI

```{r}
detect_recipe_6DAI_2020 <- recipe((treatment ~ .), 
                        data = data_2020_6dai_best_process_2020)  |>
  update_role(pot, new_role = "ID") |> 
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |>
  step_zv(all_predictors()) 

svm_wf_6DAI_2020 <- workflow() |>
    add_model(svm_spec) |>
    add_recipe(detect_recipe_6DAI_2020)

set.seed(123)
svm_fit_2020_6DAI_2020 <- svm_wf_6DAI_2020 |> fit(data_2020_6dai_best_process_2020)
```

```{r}
preds_2021_6dai <- predict(svm_fit_2020_6DAI_2020, data_2021_6dai_best_process_2020) |> 
  bind_cols(data_2021_6dai_best_process_2020$treatment) |> 
  mutate(truth = as.factor(...2))

preds_2021_6dai |> 
  accuracy(truth = truth, estimate = .pred_class)

conf_mat_2021_6dai <- preds_2021_6dai %>% 
  conf_mat(truth = truth, estimate = .pred_class) %>% 
  tidy() %>% 
  mutate(prediction = rep(c("Infected", "Control"), 2),
         truth = c(rep("Infected", 2), rep("Control", 2)),
         predicted_year = 2021,
         predicted_dai= 6)
```

### 13DAI

```{r}
detect_recipe_13DAI_2020 <- recipe((treatment ~ .), 
                        data = data_2020_13dai_best_process_2020)  |>
  update_role(pot, new_role = "ID") |> 
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |>
  step_zv(all_predictors()) 

svm_wf_13DAI_2020 <- workflow() |>
    add_model(svm_spec) |>
    add_recipe(detect_recipe_13DAI_2020)

set.seed(123)
svm_fit_2020_13DAI_2020 <- svm_wf_13DAI_2020 |> fit(data_2020_13dai_best_process_2020)
```

```{r}
preds_2021_13dai <- predict(svm_fit_2020_13DAI_2020, data_2021_13dai_best_process_2020) |> 
  bind_cols(data_2021_13dai_best_process_2020$treatment) |> 
  mutate(truth = as.factor(...2)) 

preds_2021_13dai |> 
  accuracy(truth = truth, estimate = .pred_class)

conf_mat_2021_13dai <- preds_2021_13dai %>% 
  conf_mat(truth = truth, estimate = .pred_class) %>% 
  tidy() %>% 
  mutate(prediction = rep(c("Infected", "Control"), 2),
         truth = c(rep("Infected", 2), rep("Control", 2)),
         predicted_year = 2021,
         predicted_dai= 13)
```

### 21 DAI

```{r}
detect_recipe_21DAI_2020 <- recipe((treatment ~ .), 
                        data = data_2020_21dai_best_process_2020)  |>
  update_role(pot, new_role = "ID") |> 
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |>
  step_zv(all_predictors()) 

svm_wf_21DAI_2020 <- workflow() |>
    add_model(svm_spec) |>
    add_recipe(detect_recipe_21DAI_2020)

set.seed(123)
svm_fit_2020_21DAI_2020 <- svm_wf_21DAI_2020 |> fit(data_2020_21dai_best_process_2020)
```

```{r}
preds_2021_21dai <- predict(svm_fit_2020_21DAI_2020, data_2021_21dai_best_process_2020) |> 
  bind_cols(data_2021_21dai_best_process_2020$treatment) |> 
  mutate(truth = as.factor(...2)) 

preds_2021_21dai |> 
  accuracy(truth = truth, estimate = .pred_class)

conf_mat_2021_21dai <- preds_2021_21dai %>% 
  conf_mat(truth = truth, estimate = .pred_class) %>% 
  tidy() %>% 
  mutate(prediction = rep(c("Infected", "Control"), 2),
         truth = c(rep("Infected", 2), rep("Control", 2)),
         predicted_year = 2021,
         predicted_dai= 21)
```

# Evaluate 2021 algorithms on 2020 data

## Fit final model to 2021 data (choose best one in meta_analysis_detect_ab)

## Prepare data

### 2DAI

```{r}
data_2020_2dai_best_process_2021 <- prep_dataset(dataset = hyperspec_full_2020, 
                                       alive_only = TRUE, # Not relevant here, no dead pots
                                       response_chr = "reflectance_raw", 
                                       set_date = as.Date("2020-05-14"),
                                       normalization = "Pu",
                                       VNIR_only = FALSE, 
                                       derivative = "First",
                                       scores = scores_2020)

data_2021_2dai_best_process_2021 <- prep_dataset(dataset = hyperspec_full_2021, 
                                       alive_only = TRUE, # Not relevant here, no dead pots
                                       response_chr = "reflectance_raw", 
                                       set_date = as.Date("2021-05-27"),
                                       normalization = "Pu",
                                       VNIR_only = FALSE, 
                                       derivative = "First",
                                       scores = scores_2021)

colnames(data_2021_2dai_best_process_2021) <- colnames(data_2020_2dai_best_process_2021)
```

### 6DAI

```{r}
data_2020_6dai_best_process_2021 <- prep_dataset(dataset = hyperspec_full_2020, 
                                       alive_only = TRUE, # Not relevant here, no dead pots
                                       response_chr = "reflectance_raw", 
                                       set_date = as.Date("2020-05-18"),
                                       normalization = "Pu",
                                       VNIR_only = FALSE, 
                                       derivative = "First",
                                       scores = scores_2020)

data_2021_6dai_best_process_2021 <- prep_dataset(dataset = hyperspec_full_2021, 
                                       alive_only = TRUE, # Not relevant here, no dead pots
                                       response_chr = "reflectance_raw", 
                                       set_date = as.Date("2021-05-31"),
                                       normalization = "Pu",
                                       VNIR_only = FALSE, 
                                       derivative = "First",
                                       scores = scores_2021)


colnames(data_2021_6dai_best_process_2021) <- colnames(data_2020_6dai_best_process_2021)
```

### 13DAI

```{r}
data_2020_13dai_best_process_2021 <- prep_dataset(dataset = hyperspec_full_2020, 
                                       alive_only = TRUE, # Not relevant here, no dead pots
                                       response_chr = "reflectance_raw", 
                                       set_date = as.Date("2020-05-25"),
                                       normalization = "None",
                                       VNIR_only = FALSE, 
                                       derivative = "None",
                                       scores = scores_2020)

data_2021_13dai_best_process_2021 <- prep_dataset(dataset = hyperspec_full_2021, 
                                       alive_only = TRUE, # Not relevant here, no dead pots
                                       response_chr = "reflectance_raw", 
                                       set_date = as.Date("2021-06-07"),
                                       normalization = "None",
                                       VNIR_only = FALSE, 
                                       derivative = "None",
                                       scores = scores_2021)



colnames(data_2021_13dai_best_process_2021) <- colnames(data_2020_13dai_best_process_2021)
```

### 21 DAI

```{r}
data_2020_21dai_best_process_2021 <- prep_dataset(dataset = hyperspec_full_2020, 
                                       alive_only = TRUE, # Not relevant here, no dead pots
                                       response_chr = "reflectance_raw", 
                                       set_date = as.Date("2020-06-02"),
                                       normalization = "None",
                                       VNIR_only = FALSE, 
                                       derivative = "Second",
                                       scores = scores_2020)

data_2021_21dai_best_process_2021 <- prep_dataset(dataset = hyperspec_full_2021, 
                                       alive_only = TRUE, # Not relevant here, no dead pots
                                       response_chr = "reflectance_raw", 
                                       set_date = as.Date("2021-06-07"),
                                       normalization = "None",
                                       VNIR_only = FALSE, 
                                       derivative = "Second",
                                       scores = scores_2021)



colnames(data_2021_21dai_best_process_2021) <- colnames(data_2020_21dai_best_process_2021)
```

## Train on 2021, test on 2020 

### 2DAI

```{r}
detect_recipe_2DAI_2021 <- recipe((treatment ~ .), 
                        data = data_2021_2dai_best_process_2021)  |>
  update_role(pot, new_role = "ID") |> 
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |>
  step_zv(all_predictors()) 

svm_wf_2DAI_2021 <- workflow() |>
  add_model(svm_spec) |>
  add_recipe(detect_recipe_2DAI_2021)

set.seed(123)
svm_fit_2021_2DAI_2021 <-svm_wf_2DAI_2021 |> fit(data_2021_2dai_best_process_2021)
```


```{r}
preds_2020_2dai <- predict(svm_fit_2021_2DAI_2021, data_2020_2dai_best_process_2021) |> 
  bind_cols(data_2020_2dai_best_process_2021$treatment) |> 
  mutate(truth = as.factor(...2))

preds_2020_2dai |> 
  accuracy(truth = truth, estimate = .pred_class)

conf_mat_2020_2dai <- preds_2020_2dai %>% 
  conf_mat(truth = truth, estimate = .pred_class) %>% 
  tidy() %>% 
  mutate(prediction = rep(c("Infected", "Control"), 2),
         truth = c(rep("Infected", 2), rep("Control", 2)),
         predicted_year = 2020,
         predicted_dai= 2)
```

### 6DAI

```{r}
detect_recipe_6DAI_2021 <- recipe((treatment ~ .), 
                        data = data_2021_6dai_best_process_2021)  |>
  update_role(pot, new_role = "ID") |> 
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |>
  step_zv(all_predictors()) 

svm_wf_6DAI_2021 <- workflow() |>
    add_model(svm_spec) |>
    add_recipe(detect_recipe_6DAI_2021)

set.seed(123)
svm_fit_2021_6DAI_2021 <- svm_wf_6DAI_2021 |> fit(data_2021_6dai_best_process_2021)
```

```{r}
preds_2020_6dai <- predict(svm_fit_2021_6DAI_2021, data_2020_6dai_best_process_2021) |> 
  bind_cols(data_2020_6dai_best_process_2021$treatment) |> 
  mutate(truth = as.factor(...2))

preds_2020_6dai |> 
  accuracy(truth = truth, estimate = .pred_class)

conf_mat_2020_6dai <- preds_2020_6dai %>% 
  conf_mat(truth = truth, estimate = .pred_class) %>% 
  tidy() %>% 
  mutate(prediction = rep(c("Infected", "Control"), 2),
         truth = c(rep("Infected", 2), rep("Control", 2)),
         predicted_year = 2020,
         predicted_dai= 6)
```

### 13DAI

```{r}
detect_recipe_13DAI_2021 <- recipe((treatment ~ .), 
                        data = data_2021_13dai_best_process_2021)  |>
  update_role(pot, new_role = "ID") |> 
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |>
  step_zv(all_predictors()) 

svm_wf_13DAI_2021 <- workflow() |>
    add_model(svm_spec) |>
    add_recipe(detect_recipe_13DAI_2021)

set.seed(123)
svm_fit_2021_13DAI_2021 <- svm_wf_13DAI_2021 |> fit(data_2021_13dai_best_process_2021)
```

```{r}
preds_2020_13dai <- predict(svm_fit_2021_13DAI_2021, data_2020_13dai_best_process_2021) |> 
  bind_cols(data_2020_13dai_best_process_2021$treatment) |> 
  mutate(truth = as.factor(...2)) 

preds_2020_13dai |> 
  accuracy(truth = truth, estimate = .pred_class)

conf_mat_2020_13dai <- preds_2020_13dai %>% 
  conf_mat(truth = truth, estimate = .pred_class) %>% 
  tidy() %>% 
  mutate(prediction = rep(c("Infected", "Control"), 2),
         truth = c(rep("Infected", 2), rep("Control", 2)),
         predicted_year = 2020,
         predicted_dai= 13)
```

### 21 DAI

```{r}
detect_recipe_21DAI_2021 <- recipe((treatment ~ .), 
                        data = data_2021_21dai_best_process_2021)  |>
  update_role(pot, new_role = "ID") |> 
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |>
  step_zv(all_predictors()) 

svm_wf_21DAI_2021 <- workflow() |>
    add_model(svm_spec) |>
    add_recipe(detect_recipe_21DAI_2021)

set.seed(123)
svm_fit_2021_21DAI_2021 <- svm_wf_21DAI_2021 |> fit(data_2021_21dai_best_process_2021)
```

```{r}
preds_2020_21dai <- predict(svm_fit_2021_21DAI_2021, data_2020_21dai_best_process_2021) |> 
  bind_cols(data_2020_21dai_best_process_2021$treatment) |> 
  mutate(truth = as.factor(...2)) 

preds_2020_21dai |> 
  accuracy(truth = truth, estimate = .pred_class)

conf_mat_2020_21dai <- preds_2020_21dai %>% 
  conf_mat(truth = truth, estimate = .pred_class) %>% 
  tidy() %>% 
  mutate(prediction = rep(c("Infected", "Control"), 2),
         truth = c(rep("Infected", 2), rep("Control", 2)),
         predicted_year = 2020,
         predicted_dai= 21)
```

# Combine confusion matrices

```{r}
conf_mats_pred_2020 <- rbind(conf_mat_2020_2dai, conf_mat_2020_6dai, conf_mat_2020_13dai, conf_mat_2020_21dai)
conf_mats_pred_2021 <- rbind(conf_mat_2021_2dai, conf_mat_2021_6dai, conf_mat_2021_13dai, conf_mat_2021_21dai)
```

## Predictions 2020

```{r}
conf_mats_pred_2020 %>% 
  select(-name, -predicted_year) %>% 
  group_by(predicted_dai) %>% 
  pivot_wider(names_from = truth, values_from = value) %>% 
  gt()
```


## Predictions 2021

```{r}
conf_mats_pred_2021 %>% 
  select(-name, -predicted_year) %>% 
  group_by(predicted_dai) %>% 
  pivot_wider(names_from = truth, values_from = value) %>% 
  gt()
```


```{r}
sessionInfo()
```
